Deep Reinforcement Learning for Conservation Decisions

### 2021-05-12 manuscript/manuscript.Rmd
Completing actor-critic paragraph

The general motivation is that if we know the Q function under certain conditions -- e.g. that the state-action space is discrete and that the given policy visits all state-action pairs repeatedly [@suttonbarto 131; @Qproof] --, then we can easily find the optimal action at any state by selecting the action with the highest Q-value: $a_t^* = \underset{a_t}{\text{argmax}} \, Q^\pi (s_t, a_t)$.
In the case that the state-action space can be tractably represented in a table, one could employ a  algorithm, like temporal-difference control -- see Sutton Ch. 6 for details -- which approximates the Q value for every state-action pair [@suttonbarto]. 
Yet, for many decision-making problems, the state-action space cannot be tractably represented in a table; one effective alternative for this case is to use a neural network to approximate the Q-function, a method called deep Q-learning. 
The general process for deep Q-learning differs from policy gradient methods, since instead of trying to maximize the return directly, the objective is to minimize the Bellman error.
The Bellman error is a measure of how distant a given value function is from that value function which has been re-evaluated with recent environment interaction.
In the case where the Q-function is approximated by some neural network and thus represented as $Q_\theta$, the Bellman error for one state-action pair would be

$$
  \mathcal{E} = r(s_t, a_t) + \gamma \, \underset{a}{\text{max}} \, Q_\theta(s_{t+1}, a) - Q_\theta(s_t, a_t).
$$

See the appendix/reference for details on how the Bellman Error is derived from eq. [@CITE]. 
The general scheme for approximate value-learning with deep neural networks is to sample trajectories from the environment, evaluate the Bellman Error across a range of these samples then perform the following gradient descent,

$$
  \theta_{t+1} = \theta_t - \alpha \nabla_\theta \, \sum_i \mathcal{E}_i^2,
$$

where $i$ indexes the sample.
After convergence to the Q function, the optimal policy could

There are 3 value functions that are commonly used across RL algorithms: the state-value function, $V^\pi(s_t)$, the state-action value function, $Q^\pi(s_t, a_t)$, and the advantage function, $A^\pi(s_t, a_t)$.
The state-value function is defined as the expected return from a given state,

$$
  V^\pi(s_t) = \mathbb{E}_{\tau \sim p_\pi(\tau \mid s_t)}\Bigg[\sum_{t' = t}^H \gamma^{t' - t} r(s_t, a_t)\Bigg]
$$

Meanwhile, the state-action-value or quality function, finds the expected return given a state and action,

$$
  Q^\pi(s_t, a_t) = \mathbb{E}_{\tau \sim p_\pi(\tau \mid s_t, a_t)}\Bigg[\sum_{t' = t}^H \gamma^{t' - t} r(s_t, a_t)\Bigg]
$$

And lastly, the advantage function characterizes a relative state-action value in that it quantifies how valuable a state-action pair is over a baseline.
The advantage function can then be succinctly defined as

$$
  A^\pi(s_t, a_t) = Q^\pi(s_t, a_t) - V^\pi(s_t)
$$ 
Learning $Q^\pi$, commonly called the Q-value function, is more powerful than learning the state-value function as Q-values easily prescribe what action to take when in a certain state, $a_t^* = \underset{a_t}{\text{argmax}} \, Q^\pi (s_t, a_t)$.
Conversely, if we know the state-value function, we would have to perform a maximization over an expected value to find the optimal action, $a_t^* = \underset{a_t}{\text{argmax}} \, \mathbb{E}_{s_{t+1} \sim \,T(s_{t+1} \mid s_t, a_t)}[V^\pi(s_{t+1})]$, which is less tractable.
Yet, learning the advantage function is often preferred in practice as advantage estimators favorably have lower variance than state-value and Q-value estimators [cite schulman gae].

>>>>

The general Q-learning motivation is that if we know the Q function under certain conditions -- e.g. that the state-action space is discrete and that the given policy visits all state-action pairs repeatedly [@suttonbarto 131; @Qproof] --, then we can easily find the optimal action at any state by selecting the action with the highest Q-value: $a_t^* = \underset{a_t}{\text{argmax}} \, Q^\pi (s_t, a_t)$.
In the case that the state-action space can be tractably represented in a table, one could employ a  algorithm, like temporal-difference control -- see Sutton Ch. 6 for details -- which approximates the Q value for every state-action pair [@suttonbarto]. 
Yet, for many decision-making problems, the state-action space cannot be tractably represented in a table; for this case, one effective alternative is to use a neural network to approximate the Q-function, a method called deep Q-learning. 
The general process for deep Q-learning differs from policy gradient methods, since instead of trying to maximize the return directly, the objective is to minimize the Bellman error.
The Bellman error is a measure of how distant a given value function is from that value function which has been re-evaluated with recent environment interaction.
Suppose we attempt to approximate the Q-function with a neural network and thus represent this approximate Q-function as as $Q_\phi$, the Bellman error for one state-action pair would be

$$
  \mathcal{E} = r(s_t, a_t) + \gamma \, \underset{a}{\text{max}} \, Q_\phi(s_{t+1}, a) - Q_\phi(s_t, a_t).
$$

See the appendix/reference for details on how the Bellman Error is derived from eq. [@CITE]. 
The general scheme for approximate value-learning with deep neural networks is to sample trajectories from the environment, evaluate the Bellman Error across these samples -- with the samples indexed by $i$ --, and then perform the following gradient descent:

$$
  \phi_{t+1} = \phi_t - \alpha \nabla_\phi \, \sum_i \mathcal{E}_i^2 
$$

When being evaluated, a deep Q-learning agent will select actions by inputting the current state to the Q network and then identifying the state-action pair with the highest estimated Q value.
The benefit of using neural networks to approximate Q-functions is that neural networks are able to generalize Q functions over unseen state-action pairs, whereas classic methods like TD-learning are not able to generalize [@CITE].
But there a host of problems with getting neural-network-based value-learning agents to work in practice; see @CITE for more insight on these issues.   
The primary concern with these methods is that they tend to suffer from bias which is introduced by fitting a value network to a target that incorporates an estimated value function [@CITE].

## Actor-Critic
Actor-critic algorithms integrate the main components from both value-learning and policy gradient based methods.
The "actor" attempts to learn the policy, and the "critic" attempts to learn a value function.
While the general algorithm for the critic is exactly the same as mentioned in the value-learning section, the actor differs from vanilla policy gradient algorithms by incorporating the critic in its gradient ascent step.
Instead of estimating the gradient of the return as in eq ..., the gradient of the return is estimated as

$$
  \nabla_\theta J(\pi_\theta) \approx \sum_{i,t} \nabla_\theta \log{\pi_\theta (a_{i,t} \mid s_{i,t})} \, \hat{Q}(s_{i,t}, a_{i, t}).
$$

where $\hat{Q}$ is a value function estimated by the critic.
The benefit of actor-critic algorithms is that they balance the issues with bias and variance that are common with value-learning and policy gradient methods respectively [@CITE].
Consequently, actor-critic algorithms have achieved a large majority of the state of the art results in model-free deep RL [@CITE].
